---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# corenet

<!-- badges: start -->
[![R-CMD-check](https://github.com/nptscot/corenet/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/nptscot/corenet/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

```{r, eval=FALSE, echo=FALSE}
# Code here used to set-up the package, saved for reference... 
# Check the pkg name is available
remotes::install_cran("available")
available::available("corenet")
# Create the package
usethis::use_description()
# Add package dependencies
usethis::use_package("sf")
usethis::use_package("tibble", type = "Suggests")
usethis::use_build_ignore("README.Rmd")
usethis::use_build_ignore("*.zip")
usethis::use_build_ignore("*.gpkg")
usethis::use_git_ignore("Data")
usethis::use_git_ignore("Doc")
# Add license via usethis (MIT):
usethis::use_mit_license("Zhao Wang")

# Add continuous integration
devtools::check()
usethis::use_github_action()
# Set-up website
usethis::use_pkgdown()
# Use GitHub pages:
usethis::use_github_pages()
# Action to build and push to the website with every commit:
usethis::use_pkgdown_github_pages()
# Create a function:
usethis::use_r("corenet")
# Create example osm_edinburgh_demo data object:
usethis::use_data_raw("osm_edinburgh_demo")

devtools::build_readme()
# Check the package again:
devtools::check()

# # Publish to CRAN (when ready):
# devtools::release()
```

`corenet` aims to provide a suite of functions designed to generate coherent networks for cycling, primarily for Scottish, but it can also be adapted for other regions or purposes, such as walking. 
`corenet` utilizes NPT data (www.npt.scot) as a key influencing parameter to facilitate the creation of these networks. Open road network (https://www.ordnancesurvey.co.uk/products/os-open-roads) serves as the base layer for the coherent network, chosen for its simplicity.

Install it with:

```{r, eval=FALSE}
if (!require("remotes")) {
  install.packages("remotes")
}
remotes::install_github("nptscot/corenet")
```

Load the package with the following for local development:

```{r, include=FALSE}
devtools::load_all()
```

## Minimal example

The package comes with example data for testing functions.
You can test the functions as follows:
```{r}
target_zone = zonebuilder::zb_zone("Edinburgh", n_circles = 3) |> sf::st_transform(crs = "EPSG:27700")
sf::st_crs(NPT_demo_6km) = 27700
sf::st_crs(os_edinburgh_demo_6km) = 27700
```

```{r}
OS_NPT_demo = cohesive_network_prep(base_network = os_edinburgh_demo_6km, 
                                     influence_network = NPT_demo_6km, 
                                     target_zone, 
                                     crs = "EPSG:27700", 
                                     key_attribute = "road_function", 
                                     attribute_values = c("A Road", "B Road", "Minor Road"))
#  mapview::mapview(OS_NPT_demo) + mapview::mapview(NPT_demo_6km)
```

```{r}
# The prepare_network function is invoked within the corenet function; this line allows us to examine it separately.
prepared_network = prepare_network(OS_NPT_demo, 
                                    key_attribute = "all_fastest_bicycle_go_dutch", 
                                    road_scores = list("A Road" = 1, "B Road" = 2, "Minor Road" = 1000), 
                                    transform_crs = 27700)
```

```{r}
#rename NPT_MM_OSM$geom as NPT_MM_OSM$geometry
OS_NPT_demo$geometry = OS_NPT_demo$geom

# Generate the coherent network for cycling.
CN_network = corenet(NPT_demo_6km, OS_NPT_demo, target_zone, 
                      key_attribute = "all_fastest_bicycle_go_dutch", 
                      crs = "EPSG:27700", npt_threshold = 1500, maxDistPts = 1500,
                      road_scores = list("A Road" = 1, "B Road" = 1, "Minor Road" = 100))

mapview::mapview(CN_network)
```

Group the coherent network

```{r}
grouped_network = coherent_network_group(CN_network, key_attribute = "all_fastest_bicycle_go_dutch")

# mapview::mapview(grouped_network, zcol = "group", color = viridis::viridis(12), legend = TRUE)
```

Convert the coherent network to PMtiles
```{r}
netwrok = sf::st_read("cohesive networks scotland.geojson")
create_coherent_network_PMtiles(folder_path = "", city_filename = "test", cohesive_network = grouped_network)

```


```{r}
library(dplyr)
devtools::load_all()

file_path = "inputdata/open_roads_scotland.gpkg"
open_roads_scotland = sf::read_sf(file_path)
sf::st_geometry(open_roads_scotland) = "geometry"

region = "Edinburgh and Lothians"
message("Processing coherent network for region: ", region)
region_snake = snakecase::to_snake_case(region)
cnet_path = "inputdata/combined_network_tile.geojson"
combined_net = sf::read_sf(cnet_path) |>
  sf::st_transform(crs = "EPSG:27700")

lads = sf::read_sf("inputdata/la_regions_2023.geojson")

message("Generating coherent network links for: ", region)
region_boundary = filter(lads, Region == region & LAD23NM == "City of Edinburgh") |>
  sf::st_transform(crs = "EPSG:27700")

combined_net_region_boundary = combined_net[sf::st_union(region_boundary), , op = sf::st_intersects]

min_percentile_value = stats::quantile(combined_net_region_boundary$all_fastest_bicycle_go_dutch, probs = 0.938, na.rm = TRUE)

open_roads_scotland_region_boundary = open_roads_scotland[sf::st_union(region_boundary), , op = sf::st_intersects]



OS_combined_net_region_boundary = cohesive_network_prep(
  base_network = open_roads_scotland_region_boundary,
  influence_network = combined_net_region_boundary,
  region_boundary,
  crs = "EPSG:27700",
  key_attribute = "road_function",
  attribute_values = c("A Road", "B Road", "Minor Road")
)
# 
cohesive_network_region_boundary = corenet(combined_net_region_boundary, OS_combined_net_region_boundary, region_boundary,
  key_attribute = "all_fastest_bicycle_go_dutch",
  crs = "EPSG:27700", maxDistPts = 1500, npt_threshold = min_percentile_value,
  road_scores = list("A Road" = 1, "B Road" = 2, "Minor Road" = 10),n_removeDangles = 6
)

mapview::mapview(cohesive_network_region_boundary)

cohesive_network_region_boundary = cohesive_network_region_boundary |>
  sf::st_transform(crs = "EPSG:27700")|> select(geometry)
# save cohesive_network_region_boundary as geojson
sf::st_write(cohesive_network_region_boundary, "cohesive_network_region_boundary.geojson", delete_dsn = TRUE)
mapview::mapview(cohesive_network_region_boundary) 

# read geojson and select geometry column only 
rnet_1 = sf::st_read("coherent_networks/city_of_edinburgh_2024-06-03_coherent_network.geojson") |>
  sf::st_transform(crs = "EPSG:27700") |> select(geometry)
rnet_2 = sf::st_read("coherent_networks/east_lothian_2024-06-03_coherent_network.geojson")|> 
  sf::st_transform(crs = "EPSG:27700")|> select(geometry)
rnet_3 = sf::st_read("coherent_networks/midlothian_2024-06-03_coherent_network.geojson")|>
  sf::st_transform(crs = "EPSG:27700")|> select(geometry)
rnet_4 = sf::st_read("coherent_networks/west_lothian_2024-06-03_coherent_network.geojson")|>
  sf::st_transform(crs = "EPSG:27700")|> select(geometry)

# rbind rnet_1, rnet_2, rnet_3, rnet_4 and cohesive_network_region_boundary
rnet = rbind(rnet_1, rnet_2, rnet_3, rnet_4, cohesive_network_region_boundary)
duplicated_geoms = duplicated(rnet$geometry)
rnet = rnet[!duplicated_geoms, ] |> sf::st_transform(crs = "EPSG:4326")
mapview::mapview(rnet)
# save rnet as geojson
sf::st_write(rnet, "outputdata/rnet.geojson", delete_dsn = TRUE)

  coherent_pmtiles_filename = paste0("outputdata/",region_snake, "_coherent_network",  ".pmtiles")
  
  # Construct the Tippecanoe command
  command_tippecanoe = paste0(
    'tippecanoe -o ', coherent_pmtiles_filename,
    ' --name="',  'coherent_network', '"',
    ' --layer=cohesivenetwork',
    ' --attribution="University of Leeds"',
    ' --minimum-zoom=6',
    ' --maximum-zoom=13',
    ' --maximum-tile-bytes=5000000',
    ' --simplification=10',
    ' --buffer=5',
    ' -rg',
    ' --force ',
    "outputdata/rnet.geojson"
  )
  
  # Execute the command and capture output
  system(command_tippecanoe, intern = TRUE)

values <- c(1000, 5000)
road_scores <- c(0.1, 1)

# Sensitivity parameters
a <- 2  # High sensitivity to values
b <- 2  # High sensitivity to low road_scores

# Calculate weights based on the revised formula
weights <- 1 / ((values + 1)^a * (101 - road_scores)^b) *1000000000

# Combine the data into a data frame for easy viewing
results <- data.frame(value = values, road_score = road_scores, weight = weights)
print(results)


```