---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# corenet

<!-- badges: start -->
[![R-CMD-check](https://github.com/nptscot/corenet/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/nptscot/corenet/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

```{r, eval=FALSE, echo=FALSE}
# Code here used to set-up the package, saved for reference... 
# Check the pkg name is available
remotes::install_cran("available")
available::available("corenet")
# Create the package
usethis::use_description()
# Add package dependencies
usethis::use_package("sf")
usethis::use_package("tibble", type = "Suggests")
usethis::use_build_ignore("README.Rmd")
usethis::use_build_ignore("*.zip")
usethis::use_build_ignore("*.gpkg")
usethis::use_git_ignore("Data")
usethis::use_git_ignore("Doc")
# Add license via usethis (MIT):
usethis::use_mit_license("Zhao Wang")

# Add continuous integration
devtools::check()
usethis::use_github_action()
# Set-up website
usethis::use_pkgdown()
# Use GitHub pages:
usethis::use_github_pages()
# Action to build and push to the website with every commit:
usethis::use_pkgdown_github_pages()
# Create a function:
usethis::use_r("corenet")
# Create example osm_edinburgh_demo data object:
usethis::use_data_raw("osm_edinburgh_demo")

devtools::build_readme()
# Check the package again:
devtools::check()

# # Publish to CRAN (when ready):
# devtools::release()
```

`corenet` aims to provide a suite of functions designed to generate coherent networks for cycling, primarily for Scottish, but it can also be adapted for other regions or purposes, such as walking. 
`corenet` utilizes NPT data (www.npt.scot) as a key influencing parameter to facilitate the creation of these networks. Open road network (https://www.ordnancesurvey.co.uk/products/os-open-roads) serves as the base layer for the coherent network, chosen for its simplicity.

Install it with:

```{r, eval=FALSE}
if (!require("remotes")) {
  install.packages("remotes")
}
remotes::install_github("nptscot/corenet")
```

Load the package with the following for local development:

```{r, include=FALSE}
devtools::load_all()
```

## Minimal example

The package comes with example data for testing functions.
You can test the functions as follows:
```{r}
target_zone = zonebuilder::zb_zone("Edinburgh", n_circles = 3) |> sf::st_transform(crs = "EPSG:27700")
sf::st_crs(NPT_demo_6km) = 27700
sf::st_crs(os_edinburgh_demo_6km) = 27700
```

```{r}
# Prepare the network for analysis and rename geom to geometry
os_edinburgh_demo_6km = sf::st_read("open_roads_edinburgh_6km.gpkg")
NPT_demo_6km = sf::st_read("NPT_demo_6km.gpkg")
# Rename 'geom' to 'geometry' if it is indeed named 'geom'
if ("geom" %in% names(os_edinburgh_demo_6km)) {
  os_edinburgh_demo_6km <- os_edinburgh_demo_6km %>% 
    dplyr::rename(geometry = geom)
}

if ("geom" %in% names(NPT_demo_6km)) {
  NPT_demo_6km <- NPT_demo_6km %>% 
    dplyr::rename(geometry = geom)
}
OS_NPT_demo = cohesive_network_prep(base_network = os_edinburgh_demo_6km, 
                                     influence_network = NPT_demo_6km, 
                                     target_zone, 
                                     crs = "EPSG:27700", 
                                     key_attribute = "road_function", 
                                     attribute_values = c("A Road", "B Road", "Minor Road"))
```

```{r}
# The prepare_network function is invoked within the corenet function; this line allows us to examine it separately.
prepared_network = prepare_network(OS_NPT_demo, 
                                    key_attribute = "all_fastest_bicycle_go_dutch", 
                                    road_scores = list("A Road" = 1, "B Road" = 1, "Minor Road" = 10000000), 
                                    transform_crs = 27700)
```

```{r}
#rename NPT_MM_OSM$geom as NPT_MM_OSM$geometry
OS_NPT_demo$geometry = OS_NPT_demo$geom

# Generate the coherent network for cycling.
CN_network = corenet(NPT_demo_6km, OS_NPT_demo, target_zone, 
                      key_attribute = "all_fastest_bicycle_go_dutch", 
                      crs = "EPSG:27700", dist = 10, threshold = 1500, 
                      road_scores = list("A Road" = 1, "B Road" = 1, "Minor Road" = 10000000))

# mapview::mapview(CN_network)
```

Group the coherent network

```{r}
grouped_network = coherent_network_group(CN_network, key_attribute = "all_fastest_bicycle_go_dutch")

# mapview::mapview(grouped_network, zcol = "group", color = viridis::viridis(12), legend = TRUE)
```

Convert the coherent network to PMtiles
```{r}
netwrok = sf::st_read("cohesive networks scotland.geojson")
create_coherent_network_PMtiles(folder_path = "", city_filename = "test", cohesive_network = grouped_network)

```

```{r}
library(sf)
library(dplyr)

# Read and transform initial datasets
zones_scot = sf::st_read("inputdata/las_scotland_2023.geojson") |>
  sf::st_transform(crs = "EPSG:27700")

NPT = sf::st_read("inputdata/combined_network_tile.geojson") |>
  sf::st_transform(crs = "EPSG:27700")

open_roads_national = sf::read_sf("inputdata/oproad_gb.gpkg", layer = "road_link") |>
  sf::st_transform(crs = "EPSG:27700")

# Get unique city names from LAD23NM
cities = unique(zones_scot$LAD23NM)
cities = c("West Lothian", "City of Edinburgh", "Midlothian", "East Lothian")

# Loop over each city
for (city in cities) {
  city_filename = snakecase::to_snake_case(city)
  tryCatch({
    message("Generating coherent network for: ", city)
    zone = filter(zones_scot, LAD23NM == city)

    NPT_zone = NPT[sf::st_union(zone), , op = sf::st_intersects]

    min_percentile_value = stats::quantile(NPT_zone$all_fastest_bicycle_go_dutch, probs = 0.938, na.rm = TRUE)

    open_roads_national_zone = open_roads_national[sf::st_union(zone), , op = sf::st_intersects]

    OS_NPT_zone = cohesive_network_prep(base_network = open_roads_national_zone, 
                                        influence_network = NPT_zone, 
                                        zone, 
                                        crs = "EPSG:27700", 
                                        key_attribute = "road_function", 
                                        attribute_values = c("A Road", "B Road", "Minor Road"))

    cohesive_network_zone = corenet(NPT_zone, OS_NPT_zone, zone, 
                                      key_attribute = "all_fastest_bicycle_go_dutch", 
                                      crs = "EPSG:27700", dist = 10, threshold = min_percentile_value, 
                                      road_scores = list("A Road" = 1, "B Road" = 1, "Minor Road" = 10000000))

    grouped_network = coherent_network_group(cohesive_network_zone, key_attribute = "all_fastest_bicycle_go_dutch")

    # Use city name in the filename
    create_coherent_network_PMtiles(folder_path = "outputdata/", city_filename = city_filename, cohesive_network = grouped_network)

    message("Coherent network for: ", city, " generated successfully")

  }, error = function(e) {
    message(sprintf("An error occurred with %s: %s", city, e$message))
  })  
}
```