---
format: gfm
---

# load leeds example data

```{r}
devtools::load_all()

library(sf)
library(ggplot2)
library(patchwork)
library(mapview)
library(dplyr)
library(igraph)

leeds_net = central_leeds_osm
edin_offroad_net = edinburgh_offroad |> filter(component  == 7)
```

# load two remove dangles functions
removeDangles_1
- Accepts the parameters network (an sf object) and tolerance (default = 0.001).
- Directly casts the sf object to LINESTRING using sf::st_cast
- The tolerance parameter sets a distance threshold to determine whether an endpoint is "isolated" (dangle) based on spatial proximity to other points.

removeDangles_2
- Accepts the parameters network (an sf object) and percentile (default = 0.012).
- Converts the sf object into an igraph object using sf_to_igraph, enabling the calculation of vertex degrees. This step helps identify vertices with a degree of 1, which are potential dangles.
- The percentile determines the length threshold for classifying short dangling lines for removal.

Dangle Detection:
removeDangles_1:
- Uses spatial operations to detect dangles. It extracts the endpoints of each line segment and creates a spatial buffer around them using the tolerance value.
- By finding points that do not overlap with other buffered points, it identifies isolated (dangling) endpoints. This is more of a spatial proximity-based method.

removeDangles_2:
- Uses graph theory to identify dangles. It finds vertices (nodes) with degree 1 (dangles) in the graph representation of the network.
- It then extracts the first and last coordinates of each line and checks if any endpoints are dangle vertices. This approach focuses on graph-based dangle identification.

```{r}
removeDangles_1 = function(network, tolerance = 0.001) {
    geometry_types = sf::st_geometry_type(network)
    network_linestring = network[geometry_types == "LINESTRING", ]
    network_others = network[geometry_types != "LINESTRING", ]
    network_multilinestring = network[geometry_types == "MULTILINESTRING", ]
    network_multilinestring = sf::st_cast(network_multilinestring, "LINESTRING")

    # Combine with original LINESTRING geometries
    network_lines = rbind(network_linestring, network_multilinestring)
    # Extract and combine all end points of line segments
    end_points = do.call(rbind, lapply(network_lines$geometry, function(line) {
    endpoints = rbind(sf::st_coordinates(line)[1, ], sf::st_coordinates(line)[nrow(sf::st_coordinates(line)), ])
    sf::st_as_sf(data.frame(x = endpoints[, 1], y = endpoints[, 2]), coords = c("x", "y"), crs = sf::st_crs(network))
    }))

    # Identify unique end points (potential dangles) using a spatial join to find nearby points
    buffer_points = sf::st_buffer(end_points, dist = tolerance)
    overlaps = sf::st_intersects(buffer_points, buffer_points, sparse = FALSE)
    isolated_points = end_points[rowSums(overlaps) == 1,]

    # Filter out road segments that end in these isolated points
    segments_with_dangles = sapply(sf::st_geometry(network_lines), function(geom) {
    ends = sf::st_sfc(sf::st_point(sf::st_coordinates(geom)[1,]), sf::st_point(sf::st_coordinates(geom)[nrow(sf::st_coordinates(geom)),]), crs = sf::st_crs(network))
    any(sf::st_intersects(ends, isolated_points, sparse = FALSE))
    })

    network_without_dangles = network_lines[!segments_with_dangles, ]

    return(network_without_dangles)
}

removeDangles_2 = function(network, percentile = 0.012) {

    network$length = sf::st_length(network)
    network_g = sf_to_igraph(network)
    vertex_degrees = degree(network_g)

    dangle_vertices = which(vertex_degrees == 1)

    points = unique(do.call(rbind, lapply(st_geometry(network), function(line) {
    rbind(st_coordinates(line)[1, ], st_coordinates(line)[nrow(st_coordinates(line)), ])
    })))
    # Extract indices of LINESTRINGs to check
    line_indices = sapply(st_geometry(network), function(line) {
    coords = st_coordinates(line)
    c(which(points[,1] == coords[1,1] & points[,2] == coords[1,2]),
        which(points[,1] == coords[nrow(coords),1] & points[,2] == coords[nrow(coords),2]))
    })

    # sum of network$length
    sum_value = sum(network$length)
    
    # Median of network$length
    threshold_length = percentile* sum_value
    short_dangles = network[(line_indices[1,] %in% dangle_vertices | line_indices[2,] %in% dangle_vertices) & network$length < units::set_units(threshold_length, "meters"), ]

    network_clean = network[!((line_indices[1,] %in% dangle_vertices | line_indices[2,] %in% dangle_vertices) & network$length <  units::set_units(threshold_length, "meters")), ]
}

sf_to_igraph = function(network_sf) {
  # network_sf = cycle_net_components_1
  # Assuming network_sf is LINESTRING
  points = unique(do.call(rbind, lapply(network_sf$geometry, function(line) {
    rbind(st_coordinates(line)[1, ], st_coordinates(line)[nrow(st_coordinates(line)), ])
  })))
  points = unique(points)
  edges = do.call(rbind, lapply(network_sf$geometry, function(line) {
    start = which(points[,1] == st_coordinates(line)[1,1] & points[,2] == st_coordinates(line)[1,2])
    end = which(points[,1] == st_coordinates(line)[nrow(st_coordinates(line)),1] & points[,2] == st_coordinates(line)[nrow(st_coordinates(line)),2])
    c(start, end)
  }))
  graph = graph_from_edgelist(as.matrix(edges), directed = FALSE)
  return(graph)
}

```

# compare removeDangles_1 using leeds network with different tolerance values
```{r}
RD_leeds_1 = removeDangles_1(leeds_net, tolerance = 0.001)
RD_leeds_2 = removeDangles_1(leeds_net, tolerance = 10)

# Plot for network 1
p1 <- ggplot() +
  geom_sf(data = leeds_net, color = "gray", size = 1) +
  geom_sf(data = RD_leeds_1, color = "blue", size = 1) +
  ggtitle("Network 1") +
  theme_minimal()

# Plot for network 2
p2 <- ggplot() +
  geom_sf(data = leeds_net, color = "gray", size = 1) +
  geom_sf(data = RD_leeds_2, color = "red", size = 1) +
  ggtitle("Network 2") +
  theme_minimal()

# Combine both plots side by side using patchwork
combined_plot <- p1 + p2 + plot_layout(ncol = 2)

# Display the plot
print(combined_plot)
```
# compare removeDangles_1 and removeDangles_2 using leeds network
# removeDangles_2 not applicable for leeds network
```{r}
# RD1_leeds = removeDangles_1(leeds_net, tolerance = 0.001)
# RD2_leeds = removeDangles_2(leeds_net, percentile = 0.012)

# # Plot for network 1
# p1 <- ggplot() +
#   geom_sf(data = leeds_net, color = "gray", size = 2) +
#   geom_sf(data = RD1_leeds, color = "blue", size = 1) +
#   ggtitle("Network 1") +
#   theme_minimal()

# # Plot for network 2
# p2 <- ggplot() +
#   geom_sf(data = leeds_net, color = "gray", size = 2) +
#   geom_sf(data = RD2_leeds, color = "red", size = 1) +
#   ggtitle("Network 2") +
#   theme_minimal()

# # Combine both plots side by side using patchwork
# combined_plot <- p1 + p2 + plot_layout(ncol = 2)

# # Display the plot
# print(combined_plot)
```

# compare removeDangles_1 using edinburgh offroad network with different tolerance values

# result is pretty much same for different tolerance values
```{r}
RD_edin_1 = removeDangles_1(edin_offroad_net, tolerance = 0.012)
RD_edin_2 = removeDangles_1(edin_offroad_net, tolerance = 10)

# Plot for network 1
p1 <- ggplot() +
  geom_sf(data = edin_offroad_net, color = "gray", size = 1) +
  geom_sf(data = RD_edin_1, color = "blue", size = 1) +
  ggtitle("Network using removeDangles_1_0.012") +
  theme_minimal()

# Plot for network 2
p2 <- ggplot() +
  geom_sf(data = edin_offroad_net, color = "gray", size = 1) +
  geom_sf(data = RD_edin_2, color = "red", size = 1) +
  ggtitle("using removeDangles_1_10") +
  theme_minimal()

# Combine both plots side by side using patchwork
combined_plot <- p1 + p2 + plot_layout(ncol = 2)

# Display the plot
print(combined_plot)

mapview(RD_edin_1,color = "blue") + mapview(RD_edin_2,color = "red") + mapview(edin_offroad_net, color = "black")
```

# compare removeDangles_1 and removeDangles_2 using edinburgh offroad network
removeDangles_2 is better
```{r}
RD1_edin = removeDangles_1(edin_offroad_net, tolerance = 0.001)
RD2_edin = removeDangles_2(edin_offroad_net, percentile = 0.012)

# Plot for network 1
p1 <- ggplot() +
  geom_sf(data = edin_offroad_net, color = "gray", size = 2) +
  geom_sf(data = RD1_edin, color = "blue", size = 1) +
  ggtitle("Network using removeDangles_1") +
  theme_minimal()

# Plot for network 2
p2 <- ggplot() +
  geom_sf(data = edin_offroad_net, color = "gray", size = 2) +
  geom_sf(data = RD2_edin, color = "red", size = 1) +
  ggtitle("Network using removeDangles_2") +
  theme_minimal()

# Combine both plots side by side using patchwork
combined_plot <- p1 + p2 + plot_layout(ncol = 2)

# Display the plot
print(combined_plot)

mapview(RD1_edin,color = "blue") + mapview(RD2_edin,color = "red") + mapview(edin_offroad_net, color = "black")
```

# develop a robust removeDangles function

```{r}
removeDangles_combined_refined = function(network, percentile = 0.012, tolerance = 0.001) {

  # Step 1: Cast to LINESTRING if necessary and handle MULTILINESTRING geometries
  geometry_types = sf::st_geometry_type(network)
  network_linestring = network[geometry_types == "LINESTRING", ]
  network_others = network[geometry_types != "LINESTRING", ]
  network_multilinestring = network[geometry_types == "MULTILINESTRING", ]
  network_multilinestring = suppressWarnings(sf::st_cast(network_multilinestring, "LINESTRING"))

  # Combine all LINESTRING geometries
  network_lines = rbind(network_linestring, network_multilinestring)

  # Step 2: Calculate segment lengths
  network_lines$length = sf::st_length(network_lines)

  # Step 3: Convert the network to igraph and identify dangling vertices (degree 1) and isolated vertices (degree 0)
  network_g = sf_to_igraph(network_lines)
  vertex_degrees = degree(network_g)
  
  # Isolated vertices (degree 0) and dangling vertices (degree 1)
  isolated_vertices = which(vertex_degrees == 0)
  dangle_vertices = which(vertex_degrees == 1)

  # Identify the largest connected component
  components = clusters(network_g)
  largest_component = which(components$membership == which.max(components$csize))

  # Step 4: Extract the first and last coordinates of each road segment
  points = unique(do.call(rbind, lapply(st_geometry(network_lines), function(line) {
    coords = st_coordinates(line)
    rbind(coords[1, ], coords[nrow(coords), ])
  })))

  # Ensure correct structure of edge list (start and end points)
  line_indices = sapply(st_geometry(network_lines), function(line) {
    coords = st_coordinates(line)
    c(
      which(points[, 1] == coords[1, 1] & points[, 2] == coords[1, 2]),
      which(points[, 1] == coords[nrow(coords), 1] & points[, 2] == coords[nrow(coords), 2])
    )
  })

  # Step 5: Calculate total network length and define a threshold for short dangles
  sum_value = sum(network_lines$length)
  threshold_length = percentile * sum_value

  # Step 6: Spatial proximity check (from removeDangles_1)
  end_points = do.call(rbind, lapply(network_lines$geometry, function(line) {
    endpoints = rbind(sf::st_coordinates(line)[1, ], sf::st_coordinates(line)[nrow(sf::st_coordinates(line)), ])
    sf::st_as_sf(data.frame(x = endpoints[, 1], y = endpoints[, 2]), coords = c("x", "y"), crs = sf::st_crs(network))
  }))

  # Buffer endpoints and check for overlaps
  buffer_points = sf::st_buffer(end_points, dist = tolerance)
  overlaps = sf::st_intersects(buffer_points, buffer_points, sparse = FALSE)
  spatially_isolated_points = end_points[rowSums(overlaps) == 1, ]

  # Step 7: Remove segments that are short dangles or connected to isolated points (degree 0 or 1)
  segments_with_dangles_or_isolated = sapply(seq_len(length(st_geometry(network_lines))), function(i) {
    geom = st_geometry(network_lines)[i]
    ends = sf::st_sfc(
      sf::st_point(sf::st_coordinates(geom)[1, ]), 
      sf::st_point(sf::st_coordinates(geom)[nrow(sf::st_coordinates(geom)), ]), 
      crs = sf::st_crs(network)
    )
    
    # Check if the endpoints are in isolated vertices (degree 0), dangle vertices (degree 1), or spatially isolated points
    isolated_endpoints = sf::st_intersects(ends, spatially_isolated_points, sparse = FALSE)
    
    # Convert to single TRUE or FALSE for isolated points
    is_isolated = any(rowSums(isolated_endpoints) > 0)
    
    # Check if the segment is connected to isolated or dangling vertices
    is_dangle_or_isolated_vertex = any(line_indices[1, i] %in% isolated_vertices | line_indices[2, i] %in% isolated_vertices) ||
      any(line_indices[1, i] %in% dangle_vertices | line_indices[2, i] %in% dangle_vertices)
    
    # Check if the segment is a short dangle
    is_short_dangle = (line_indices[1, i] %in% dangle_vertices | line_indices[2, i] %in% dangle_vertices) & 
      network_lines$length[i] < units::set_units(threshold_length, "meters")
    
    # Check if the segment is part of the largest connected component
    in_largest_component = any(line_indices[1, i] %in% largest_component | line_indices[2, i] %in% largest_component)

    # Return TRUE if any of the conditions are met, but exclude segments in the largest component
    return((is_isolated || is_dangle_or_isolated_vertex || is_short_dangle) && !in_largest_component)
  })

  # Remove the identified segments
  network_clean = network_lines[!segments_with_dangles_or_isolated, ]

  return(network_clean)
}
```

```{r}
new_RD_leeds = removeDangles_combined(leeds_net, percentile = 0.0001, tolerance = 0.0001)
new_RD_edin = removeDangles_combined(edin_offroad_net, percentile = 0.0001, tolerance = 0.0001)

mapview(leeds_net, color = "black") + mapview(new_RD_leeds, color = "red") 
mapview(edin_offroad_net, color = "black") + mapview(new_RD_edin, color = "red")
```

```{r}
remove_dangles_by_degree = function(network) {

  # Step 1: Cast to LINESTRING if necessary and handle MULTILINESTRING geometries
  geometry_types = sf::st_geometry_type(network)
  network_linestring = network[geometry_types == "LINESTRING", ]
  network_others = network[geometry_types != "LINESTRING", ]
  network_multilinestring = network[geometry_types == "MULTILINESTRING", ]
  network_multilinestring = suppressWarnings(sf::st_cast(network_multilinestring, "LINESTRING"))

  # Combine all LINESTRING geometries
  network_lines = rbind(network_linestring, network_multilinestring)

  # Step 2: Convert the network to igraph and identify dangling (degree 1) and isolated (degree 0) vertices
  network_g = sf_to_igraph(network_lines)
  vertex_degrees = degree(network_g)

  # Isolated vertices (degree 0) and dangling vertices (degree 1)
  isolated_vertices = which(vertex_degrees == 0)
  dangle_vertices = which(vertex_degrees == 1)

  # Step 3: Extract the first and last coordinates of each road segment
  points = unique(do.call(rbind, lapply(st_geometry(network_lines), function(line) {
    coords = st_coordinates(line)
    rbind(coords[1, ], coords[nrow(coords), ])
  })))

  # Ensure correct structure of edge list (start and end points)
  line_indices = sapply(st_geometry(network_lines), function(line) {
    coords = st_coordinates(line)
    c(
      which(points[, 1] == coords[1, 1] & points[, 2] == coords[1, 2]),
      which(points[, 1] == coords[nrow(coords), 1] & points[, 2] == coords[nrow(coords), 2])
    )
  })

  # Step 4: Remove segments that are connected to vertices with degree 1 or 0
  segments_to_remove = sapply(seq_len(length(st_geometry(network_lines))), function(i) {
    # Check if the segment's endpoints are connected to isolated (degree 0) or dangling (degree 1) vertices
    return(any(line_indices[1, i] %in% isolated_vertices | line_indices[2, i] %in% isolated_vertices) ||
           any(line_indices[1, i] %in% dangle_vertices | line_indices[2, i] %in% dangle_vertices))
  })

  # Keep only the segments that are NOT connected to vertices with degree 1 or 0
  network_clean = network_lines[!segments_to_remove, ]

  return(network_clean)
}
```

```{r}
new_RD_leeds = remove_dangles_by_degree(combined_network)
new_RD_edin = remove_dangles_by_degree(edin_offroad_net)

mapview(leeds_net, color = "black") + mapview(new_RD_leeds, color = "red")
mapview(edin_offroad_net, color = "black") + mapview(new_RD_edin, color = "red")
```

```{r}
library(sf)
library(units)
# Step 1: Extract endpoints of each line segment
endpoints <- do.call(rbind, lapply(st_geometry(combined_network), function(line) {
  coords <- st_coordinates(line)
  rbind(coords[1, ], coords[nrow(coords), ])
}))

# Create an sf object for the endpoints
endpoints_sf <- st_as_sf(data.frame(x = endpoints[, 1], y = endpoints[, 2]), coords = c("x", "y"), crs = st_crs(combined_network))

# Step 2: Calculate distances between endpoints of different lines
distance_matrix <- st_distance(endpoints_sf)

# Step 3: Set a threshold with proper units (e.g., meters)
# Use the same units as the distance_matrix (units should match)
threshold <- set_units(0.001, "m")  # Adjust 0.001 as needed

# Step 4: Detect gaps: Find where the distance between endpoints is larger than the threshold
gap_indices <- which(as.numeric(distance_matrix) > as.numeric(threshold) & as.numeric(distance_matrix) != 0, arr.ind = TRUE)

# Check if gap_indices is a vector or matrix
if (is.matrix(gap_indices)) {
  gap_points <- endpoints_sf[gap_indices[, 1], ]
} else if (is.vector(gap_indices)) {
  gap_points <- endpoints_sf[gap_indices, ]
} else {
  stop("Unexpected format for gap_indices.")
}

# Plot the network and gaps
ggplot() +
  geom_sf(data = leeds_net, color = "black", size = 0.5) +    # Plot the network lines
  geom_sf(data = gap_points, color = "red", size = 3) +     # Plot the gap points
  ggtitle("Network Gaps") +
  theme_minimal()


```

```{r}
library(sf)
library(units)

network = leeds_net

# Step 1: Extract endpoints of each line segment
endpoints <- do.call(rbind, lapply(st_geometry(network), function(line) {
  coords <- st_coordinates(line)
  rbind(coords[1, ], coords[nrow(coords), ])
}))

# Create an sf object for the endpoints
endpoints_sf <- st_as_sf(data.frame(x = endpoints[, 1], y = endpoints[, 2]), coords = c("x", "y"), crs = st_crs(network))

# Ensure that the geometry is cast to POINT type
endpoints_sf <- st_cast(endpoints_sf, "POINT")

# Step 2: Calculate distances between endpoints of different lines
distance_matrix <- st_distance(endpoints_sf)

# Step 3: Set a threshold for closing the gaps (e.g., 0.01 meters)
threshold <- set_units(0.01, "m")

# Detect gaps: Find where the distance between endpoints is larger than 0 and less than the threshold
gap_indices <- which(as.numeric(distance_matrix) > 0 & as.numeric(distance_matrix) <= as.numeric(threshold), arr.ind = TRUE)

# Check if gaps are detected
if (length(gap_indices) == 0) {
  stop("No gaps detected within the specified threshold.")
}

# Ensure gap_indices is a matrix
if (!is.matrix(gap_indices)) {
  gap_indices <- matrix(gap_indices, ncol = 2)
}

# Step 5: Create new LINESTRING segments to close the gaps
new_lines <- list()
for (i in seq_len(nrow(gap_indices))) {
  # Extract coordinates directly from endpoints_sf using st_coordinates
  p1 <- st_coordinates(endpoints_sf[gap_indices[i, 1], ])
  p2 <- st_coordinates(endpoints_sf[gap_indices[i, 2], ])
  
  # Create a new LINESTRING connecting the two points
  new_line <- st_linestring(rbind(p1, p2))
  new_lines[[i]] <- new_line
}

# Step 6: Convert the new lines to an sf object
new_lines_sf <- st_sfc(new_lines, crs = st_crs(network))

# Step 7: Combine the original network with the new line segments
combined_network <- rbind(network, st_as_sf(new_lines_sf))

# Step 8: Visualize the updated network
ggplot() +
  geom_sf(data = combined_network, color = "black", size = 0.5) +   # Plot the combined network (including new lines)
  ggtitle("Network with Closed Gaps") +
  theme_minimal()

```
