---
format: gfm
---

# load leeds example data

```{r}
devtools::load_all()

library(sf)
library(ggplot2)
library(patchwork)
library(mapview)
library(dplyr)
library(igraph)

leeds_net = central_leeds_osm
edin_offroad_net = edinburgh_offroad |> filter(component  == 7)
```

# load two remove dangles functions
removeDangles_1
- Accepts the parameters network (an sf object) and tolerance (default = 0.001).
- Directly casts the sf object to LINESTRING using sf::st_cast
- The tolerance parameter sets a distance threshold to determine whether an endpoint is "isolated" (dangle) based on spatial proximity to other points.

removeDangles_2
- Accepts the parameters network (an sf object) and percentile (default = 0.012).
- Converts the sf object into an igraph object using sf_to_igraph, enabling the calculation of vertex degrees. This step helps identify vertices with a degree of 1, which are potential dangles.
- The percentile determines the length threshold for classifying short dangling lines for removal.

Dangle Detection:
removeDangles_1:
- Uses spatial operations to detect dangles. It extracts the endpoints of each line segment and creates a spatial buffer around them using the tolerance value.
- By finding points that do not overlap with other buffered points, it identifies isolated (dangling) endpoints. This is more of a spatial proximity-based method.

removeDangles_2:
- Uses graph theory to identify dangles. It finds vertices (nodes) with degree 1 (dangles) in the graph representation of the network.
- It then extracts the first and last coordinates of each line and checks if any endpoints are dangle vertices. This approach focuses on graph-based dangle identification.

```{r}
removeDangles_1 = function(network, tolerance = 0.001) {
    geometry_types = sf::st_geometry_type(network)
    network_linestring = network[geometry_types == "LINESTRING", ]
    network_others = network[geometry_types != "LINESTRING", ]
    network_multilinestring = network[geometry_types == "MULTILINESTRING", ]
    network_multilinestring = sf::st_cast(network_multilinestring, "LINESTRING")

    # Combine with original LINESTRING geometries
    network_lines = rbind(network_linestring, network_multilinestring)
    # Extract and combine all end points of line segments
    end_points = do.call(rbind, lapply(network_lines$geometry, function(line) {
    endpoints = rbind(sf::st_coordinates(line)[1, ], sf::st_coordinates(line)[nrow(sf::st_coordinates(line)), ])
    sf::st_as_sf(data.frame(x = endpoints[, 1], y = endpoints[, 2]), coords = c("x", "y"), crs = sf::st_crs(network))
    }))

    # Identify unique end points (potential dangles) using a spatial join to find nearby points
    buffer_points = sf::st_buffer(end_points, dist = tolerance)
    overlaps = sf::st_intersects(buffer_points, buffer_points, sparse = FALSE)
    isolated_points = end_points[rowSums(overlaps) == 1,]

    # Filter out road segments that end in these isolated points
    segments_with_dangles = sapply(sf::st_geometry(network_lines), function(geom) {
    ends = sf::st_sfc(sf::st_point(sf::st_coordinates(geom)[1,]), sf::st_point(sf::st_coordinates(geom)[nrow(sf::st_coordinates(geom)),]), crs = sf::st_crs(network))
    any(sf::st_intersects(ends, isolated_points, sparse = FALSE))
    })

    network_without_dangles = network_lines[!segments_with_dangles, ]

    return(network_without_dangles)
}

removeDangles_2 = function(network, percentile = 0.012) {

    network$length = sf::st_length(network)
    network_g = sf_to_igraph(network)
    vertex_degrees = degree(network_g)

    dangle_vertices = which(vertex_degrees == 1)

    points = unique(do.call(rbind, lapply(st_geometry(network), function(line) {
    rbind(st_coordinates(line)[1, ], st_coordinates(line)[nrow(st_coordinates(line)), ])
    })))
    # Extract indices of LINESTRINGs to check
    line_indices = sapply(st_geometry(network), function(line) {
    coords = st_coordinates(line)
    c(which(points[,1] == coords[1,1] & points[,2] == coords[1,2]),
        which(points[,1] == coords[nrow(coords),1] & points[,2] == coords[nrow(coords),2]))
    })

    # sum of network$length
    sum_value = sum(network$length)
    
    # Median of network$length
    threshold_length = percentile* sum_value
    short_dangles = network[(line_indices[1,] %in% dangle_vertices | line_indices[2,] %in% dangle_vertices) & network$length < units::set_units(threshold_length, "meters"), ]

    network_clean = network[!((line_indices[1,] %in% dangle_vertices | line_indices[2,] %in% dangle_vertices) & network$length <  units::set_units(threshold_length, "meters")), ]
}

sf_to_igraph = function(network_sf) {
  # network_sf = cycle_net_components_1
  # Assuming network_sf is LINESTRING
  points = unique(do.call(rbind, lapply(network_sf$geometry, function(line) {
    rbind(st_coordinates(line)[1, ], st_coordinates(line)[nrow(st_coordinates(line)), ])
  })))
  points = unique(points)
  edges = do.call(rbind, lapply(network_sf$geometry, function(line) {
    start = which(points[,1] == st_coordinates(line)[1,1] & points[,2] == st_coordinates(line)[1,2])
    end = which(points[,1] == st_coordinates(line)[nrow(st_coordinates(line)),1] & points[,2] == st_coordinates(line)[nrow(st_coordinates(line)),2])
    c(start, end)
  }))
  graph = graph_from_edgelist(as.matrix(edges), directed = FALSE)
  return(graph)
}

```

# compare removeDangles_1 using leeds network with different tolerance values
```{r}
RD_leeds_1 = removeDangles_1(leeds_net, tolerance = 0.001)
RD_leeds_2 = removeDangles_1(leeds_net, tolerance = 10)

# Plot for network 1
p1 = ggplot() +
  geom_sf(data = leeds_net, color = "gray", size = 1) +
  geom_sf(data = RD_leeds_1, color = "blue", size = 1) +
  ggtitle("Network 1") +
  theme_minimal()

# Plot for network 2
p2 = ggplot() +
  geom_sf(data = leeds_net, color = "gray", size = 1) +
  geom_sf(data = RD_leeds_2, color = "red", size = 1) +
  ggtitle("Network 2") +
  theme_minimal()

# Combine both plots side by side using patchwork
combined_plot = p1 + p2 + plot_layout(ncol = 2)

# Display the plot
print(combined_plot)
```
# compare removeDangles_1 and removeDangles_2 using leeds network
# removeDangles_2 not applicable for leeds network
```{r}
# RD1_leeds = removeDangles_1(leeds_net, tolerance = 0.001)
# RD2_leeds = removeDangles_2(leeds_net, percentile = 0.012)

# # Plot for network 1
# p1 = ggplot() +
#   geom_sf(data = leeds_net, color = "gray", size = 2) +
#   geom_sf(data = RD1_leeds, color = "blue", size = 1) +
#   ggtitle("Network 1") +
#   theme_minimal()

# # Plot for network 2
# p2 = ggplot() +
#   geom_sf(data = leeds_net, color = "gray", size = 2) +
#   geom_sf(data = RD2_leeds, color = "red", size = 1) +
#   ggtitle("Network 2") +
#   theme_minimal()

# # Combine both plots side by side using patchwork
# combined_plot = p1 + p2 + plot_layout(ncol = 2)

# # Display the plot
# print(combined_plot)
```

# compare removeDangles_1 using edinburgh offroad network with different tolerance values

# result is pretty much same for different tolerance values
```{r}
RD_edin_1 = removeDangles_1(edin_offroad_net, tolerance = 0.012)
RD_edin_2 = removeDangles_1(edin_offroad_net, tolerance = 10)

# Plot for network 1
p1 = ggplot() +
  geom_sf(data = edin_offroad_net, color = "gray", size = 1) +
  geom_sf(data = RD_edin_1, color = "blue", size = 1) +
  ggtitle("Network using removeDangles_1_0.012") +
  theme_minimal()

# Plot for network 2
p2 = ggplot() +
  geom_sf(data = edin_offroad_net, color = "gray", size = 1) +
  geom_sf(data = RD_edin_2, color = "red", size = 1) +
  ggtitle("using removeDangles_1_10") +
  theme_minimal()

# Combine both plots side by side using patchwork
combined_plot = p1 + p2 + plot_layout(ncol = 2)

# Display the plot
print(combined_plot)

mapview(RD_edin_1,color = "blue") + mapview(RD_edin_2,color = "red") + mapview(edin_offroad_net, color = "black")
```

# compare removeDangles_1 and removeDangles_2 using edinburgh offroad network
removeDangles_2 is better
```{r}
RD1_edin = removeDangles_1(edin_offroad_net, tolerance = 0.001)
RD2_edin = removeDangles_2(edin_offroad_net, percentile = 0.012)

# Plot for network 1
p1 = ggplot() +
  geom_sf(data = edin_offroad_net, color = "gray", size = 2) +
  geom_sf(data = RD1_edin, color = "blue", size = 1) +
  ggtitle("Network using removeDangles_1") +
  theme_minimal()

# Plot for network 2
p2 = ggplot() +
  geom_sf(data = edin_offroad_net, color = "gray", size = 2) +
  geom_sf(data = RD2_edin, color = "red", size = 1) +
  ggtitle("Network using removeDangles_2") +
  theme_minimal()

# Combine both plots side by side using patchwork
combined_plot = p1 + p2 + plot_layout(ncol = 2)

# Display the plot
print(combined_plot)

mapview(RD1_edin,color = "blue") + mapview(RD2_edin,color = "red") + mapview(edin_offroad_net, color = "black")
```

# display the node degree of the network
```{r}
# Load necessary libraries
library(sf)
library(igraph)
library(dplyr)
library(ggplot2)
library(tidyr)  # Load tidyr for the separate function

# Extract edges (line segments)
edges = st_cast(leeds_net, "LINESTRING")
fix_edges = st_snap(edges |> st_transform(27700), st_union(edges) |> st_transform(27700), tolerance = 0.1)
edges = fix_edges
# Extract start and end coordinates of each line segment
edge_list = st_coordinates(edges) %>%
  as.data.frame() %>%
  group_by(L1) %>%
  summarize(
    x_start = first(X),
    y_start = first(Y),
    x_end = last(X),
    y_end = last(Y)
  )
# Create node identifiers by concatenating coordinates
edges_df = edge_list %>%
  mutate(
    from = paste(x_start, y_start, sep = ","),
    to = paste(x_end, y_end, sep = ",")
  ) %>%
  select(from, to)
# Create an undirected graph
g = graph_from_data_frame(edges_df, directed = FALSE)

node_degree = degree(g)

# Convert node_degree to a data frame
node_degree_df = data.frame(
  node = names(node_degree),
  degree = node_degree
)

# Split the node coordinates back into separate columns
node_degree_df = node_degree_df %>%
  separate(node, into = c("X", "Y"), sep = ",", convert = TRUE)

# Convert node_degree_df back to sf object
node_degree_sf = st_as_sf(node_degree_df, coords = c("X", "Y"), crs = st_crs(27700))

mapview(node_degree_sf |> st_transform(4326), zcol = "degree") + mapview(leeds_net)


plot_networks = function(networks, colors, titles, output_file, ncol = 2, width = 12, height = 6, base_map = TRUE, line_width = 1, point_size = 1) {
  
  plots = list()  # Create a list to hold individual plots

  for (i in seq_along(networks)) {
    network_plot = ggplot()  # Initialize ggplot

    if (base_map) {
      network_plot = network_plot + annotation_map_tile(type = "osm")  # Add OSM base map if enabled
    }

    # Loop through each network data set
    if (is.list(networks[[i]]) && length(networks[[i]]) == 2) {
      for (j in 1:2) {
        data = networks[[i]][[j]]
        geom_type = sf::st_geometry_type(data)[1]
        size_use = if (geom_type %in% c("POINT", "MULTIPOINT")) point_size else line_width
        
        if (is.character(colors[[i]][j]) && any(colors[[i]][j] %in% names(data))) {
          network_plot = network_plot +
            geom_sf(data = data, aes(color = factor(.data[[colors[[i]][j]]])), size = size_use)
        } else {
          network_plot = network_plot +
            geom_sf(data = data, color = colors[[i]][j], size = size_use)
        }
      }
    } else {
      data = networks[[i]]
      geom_type = sf::st_geometry_type(data)[1]
      size_use = if (geom_type %in% c("POINT", "MULTIPOINT")) point_size else line_width

      if (is.character(colors[[i]]) && any(colors[[i]] %in% names(data))) {
        network_plot = network_plot +
          geom_sf(data = data, aes(color = factor(.data[[colors[[i]]]])), size = size_use)
      } else {
        network_plot = network_plot +
          geom_sf(data = data, color = colors[[i]], size = size_use)
      }
    }
    
    network_plot = network_plot +
      ggtitle(titles[[i]]) +
      theme_minimal() +
      theme(panel.background = element_rect(fill = "gainsboro"), legend.position = "right") +
      scale_color_viridis(discrete = TRUE)  # Apply discrete color scale

    plots[[i]] = network_plot
  }

  combined_plot = patchwork::wrap_plots(plots, ncol = ncol)
  ggsave(output_file, combined_plot, width = width, height = height)
  
  return(combined_plot)
}

# Example usage
# Assuming leeds_net and node_degree_sf are properly defined sf objects
networks = list(
  list(leeds_net, node_degree_sf)
)
colors = list(
  c("blue", "degree")
)
titles = list(
  "Network Visualization"
)

# Call the function
plot_networks(networks, colors, titles, "networks_plot.png", ncol = 1, base_map = FALSE, line_width = 2, point_size = 5)

```


# new remove dangles function
```{r}
remove_dangles = function(network, crs_proj = 27700, tolerance = 1, iterative = TRUE) {

  # Step 1: Transform to Projected CRS
  projected_crs = sf::st_crs(crs_proj)
  network_proj = sf::st_transform(network, crs = projected_crs)
  
  # Step 2: Node the Network (Split Lines at Intersections)
  network_noded = sf::st_union(network_proj)
  network_noded = sf::st_line_merge(network_noded)
  network_noded_sf = sf::st_cast(network_noded, "LINESTRING")
  
  # Step 3: Snap the Network to Itself 
  threshold = units::set_units(tolerance, "m")
  network_snapped = sf::st_snap(network_noded_sf, sf::st_union(network_noded_sf), tolerance = threshold)
  
  # Step 4: Extract Edges (Line Segments)
  edges = sf::st_cast(network_snapped, "LINESTRING")
  
  # Step 5: Create Edge List with Start and End Coordinates
  edge_list = sf::st_coordinates(edges) |>
    as.data.frame() |>
    dplyr::group_by(L1) |>
    dplyr::summarize(
      x_start = dplyr::first(X),
      y_start = dplyr::first(Y),
      x_end = dplyr::last(X),
      y_end = dplyr::last(Y)
    )
  
  # Step 6: Create Node Identifiers
  edges_df = edge_list |>
    dplyr::mutate(
      from = paste(x_start, y_start, sep = ","),
      to = paste(x_end, y_end, sep = ",")
    ) |>
    dplyr::select(from, to)
  
  # Step 7: Create Graph
  g = igraph::graph_from_data_frame(edges_df, directed = FALSE)
  
  # Step 8: Remove Dangles (Iteratively or Once)
  if (iterative) {
    # Iterative Removal of Dangles
    repeat {
      # Calculate Node Degrees
      node_degree = igraph::degree(g)
      
      # Break the loop if no nodes with degree 1 are left
      if (all(node_degree > 1)) {
        break
      }
      
      # Identify Nodes with Degree 1
      node_degree_df = data.frame(
        node = names(node_degree),
        degree = node_degree
      ) |>
        tidyr::separate(node, into = c("X", "Y"), sep = ",", convert = TRUE)
      
      # Identify Dangle Nodes
      dangle_nodes = node_degree_df |> dplyr::filter(degree == 1)
      
      # Create Node Info for Joining
      node_info = node_degree_df
      
      # Merge Edge List with Node Degrees to Identify Edges Connected to Dangle Nodes
      edges_with_nodes = edge_list |>
        dplyr::mutate(
          from = paste(x_start, y_start, sep = ","),
          to = paste(x_end, y_end, sep = ",")
        ) |>
        dplyr::left_join(
          node_info |>
            dplyr::mutate(node = paste(X, Y, sep = ",")),
          by = c("from" = "node")
        ) |>
        dplyr::rename(degree_from = degree) |>
        dplyr::left_join(
          node_info |>
            dplyr::mutate(node = paste(X, Y, sep = ",")),
          by = c("to" = "node")
        ) |>
        dplyr::rename(degree_to = degree)
      
      # Identify Edges Connected to Dangle Nodes
      dangle_edges_indices = edges_with_nodes |>
        dplyr::filter(degree_from == 1 | degree_to == 1) |>
        dplyr::pull(L1)
      
      # Break if No Dangle Edges are Found
      if (length(dangle_edges_indices) == 0) {
        break
      }
      
      # Remove Dangle Edges
      edges = edges[-dangle_edges_indices, ]
      
      # Reconstruct Edge List and Graph for Next Iteration
      edge_list = sf::st_coordinates(edges) |>
        as.data.frame() |>
        dplyr::group_by(L1) |>
        dplyr::summarize(
          x_start = dplyr::first(X),
          y_start = dplyr::first(Y),
          x_end = dplyr::last(X),
          y_end = dplyr::last(Y)
        )
      
      edges_df = edge_list |>
        dplyr::mutate(
          from = paste(x_start, y_start, sep = ","),
          to = paste(x_end, y_end, sep = ",")
        ) |>
        dplyr::select(from, to)
      
      g = igraph::graph_from_data_frame(edges_df, directed = FALSE)
    }
  } else {
    # Single Removal of Dangles
    node_degree = igraph::degree(g)
    
    node_degree_df = data.frame(
      node = names(node_degree),
      degree = node_degree
    ) |>
      tidyr::separate(node, into = c("X", "Y"), sep = ",", convert = TRUE)

    # Convert node_degree_df back to sf object
    node_degree_sf = sf::st_as_sf(node_degree_df, coords = c("X", "Y"), crs = st_crs(crs_proj))

    node_info = node_degree_df

    edges_with_nodes = edge_list |>
      dplyr::mutate(
        from = paste(x_start, y_start, sep = ","),
        to = paste(x_end, y_end, sep = ",")
      ) |>
      dplyr::left_join(
        node_info |>
          dplyr::mutate(node = paste(X, Y, sep = ",")),
        by = c("from" = "node")
      ) |>
      dplyr::rename(degree_from = degree) |>
      dplyr::left_join(
        node_info |>
          dplyr::mutate(node = paste(X, Y, sep = ",")),
        by = c("to" = "node")
      ) |>
      dplyr::rename(degree_to = degree)
    
    dangle_edges_indices = edges_with_nodes |>
      dplyr::filter(degree_from == 1 | degree_to == 1) |>
      dplyr::pull(L1)
   
    edges = edges[-dangle_edges_indices, ]
  }
  
  # Step 9: Return the Cleaned Network
  cleaned_network = sf::st_transform(edges, sf::st_crs(network))
  cleaned_network = sf::st_sf(cleaned_network)
  return(cleaned_network)
}

```

# apply to leeds network and edinburgh offroad network

```{r}
cleaned_leeds = remove_dangles(leeds_net, crs_proj = 27700, tolerance = 1, iterative = FALSE) |> st_transform(27700)
leeds_net = leeds_net |> st_transform(27700)

cleaned_edin = remove_dangles(edin_offroad_net, crs_proj = 27700, tolerance = 1,iterative = FALSE)
names(cleaned_edin)

mapview(cleaned_leeds, color = "red") + mapview(leeds_net, color = "blue")

mapview(cleaned_edin, color = "red") + mapview(edin_offroad_net, color = "blue")
```
```{r}
remove_dangles_with_angle = function(network, crs_proj = 27700, tolerance = 1,
                                      angle_threshold_low = 10, angle_threshold_high = 170,
                                      return_debug_data = FALSE) {
  # Load required packages
  library(sf)
  library(dplyr)
  library(tidyr)
  library(igraph)
  library(units)
  
  # Step 1: Transform to Projected CRS
  projected_crs = st_crs(crs_proj)
  network_proj = st_transform(network, crs = projected_crs)
  
  # Step 2: Node the Network (Split Lines at Intersections)
  network_noded = st_union(network_proj)
  network_noded_sf = st_line_merge(network_noded)
  network_noded_sf = st_cast(network_noded_sf, "LINESTRING")
  
  # Step 3: Snap the Network to Itself
  threshold = set_units(tolerance, "m")
  network_snapped = st_snap(network_noded_sf, st_union(network_noded_sf), tolerance = threshold)
  
  # Step 4: Extract Edges (Line Segments)
  edges = st_cast(network_snapped, "LINESTRING")
  
  # Ensure 'edges' is an sf object with an ID column
  edges = st_sf(geometry = edges)
  edges$edge_id = seq_len(nrow(edges))
  
  # Step 5: Create Edge List with Start and End Coordinates
  edge_coords = st_coordinates(edges)
  edges_coords_df = as.data.frame(edge_coords)
  edges_coords_df$edge_id = edges$edge_id[edges_coords_df$L1]
  
  # Group by edge and get start and end points
  edges_with_nodes = edges_coords_df %>%
    group_by(edge_id) %>%
    summarize(
      x_start = first(X),
      y_start = first(Y),
      x_end = last(X),
      y_end = last(Y)
    ) %>%
    ungroup()
  
  # Step 6: Create Node Identifiers
  edges_with_nodes = edges_with_nodes %>%
    mutate(
      from = paste(x_start, y_start, sep = ","),
      to = paste(x_end, y_end, sep = ",")
    )
  
  edges_df = edges_with_nodes %>%
    select(edge_id, from, to)
  
  # Step 7: Create Graph
  g = graph_from_data_frame(edges_df %>% select(from, to), directed = FALSE)
  
  # Step 8: Calculate Node Degrees
  node_degree = degree(g)
  
  # Create Node Data Frame
  nodes_df = data.frame(
    node = names(node_degree),
    degree = node_degree
  ) %>%
    separate(node, into = c("X", "Y"), sep = ",", convert = TRUE, remove = FALSE) %>%
    mutate(node_id = row_number())
  
  # Map Edges to Node IDs and Degrees
  edges_with_nodes = edges_with_nodes %>%
    left_join(
      nodes_df %>% select(node, node_id, degree) %>% rename(degree_from = degree),
      by = c("from" = "node")
    ) %>%
    rename(from_node_id = node_id) %>%
    left_join(
      nodes_df %>% select(node, node_id, degree) %>% rename(degree_to = degree),
      by = c("to" = "node")
    ) %>%
    rename(to_node_id = node_id)
  
  # Step 9: Identify Candidate Edges Based on Degree
  candidate_edges = edges_with_nodes %>%
    filter(
      degree_from >= 3 & degree_to == 1
    )
  
  # Initialize columns in candidate_edges
  candidate_edges$min_angle_deg = NA_real_
  candidate_edges$is_dangle = FALSE
  
  # Include the corrected calculate_angle_at_node function here
  # (Provided above)
  
  # Initialize vector to store dangle edge IDs
  dangle_edge_ids = c()
  
  # Loop over candidate edges
  for (i in seq_len(nrow(candidate_edges))) {
    edge = candidate_edges[i, ]
    candidate_edge_id = edge$edge_id
    
    # The branch node is the from_node (degree >= 3)
    branch_node_id = edge$from_node_id
    
    # Get edges connected to the branch node, excluding the candidate edge
    connected_edges = edges_with_nodes %>%
      filter(
        (from_node_id == branch_node_id | to_node_id == branch_node_id) &
          edge_id != candidate_edge_id
      )
    
    # Initialize min_angle_deg for this candidate edge
    min_angle = NA_real_
    
    # Initialize flag to determine if edge is continuous
    is_continuous = FALSE
    
    # Calculate angle between candidate edge and each connected edge
    for (j in seq_len(nrow(connected_edges))) {
      other_edge_id = connected_edges$edge_id[j]
      angle_deg = calculate_angle_at_node(candidate_edge_id, other_edge_id,
                                           branch_node_id, edges_with_nodes, nodes_df)
      
      # Update min_angle
      if (!is.na(angle_deg)) {
        if (is.na(min_angle) || angle_deg < min_angle) {
          min_angle = angle_deg
        }
      }
      
      # Check if angle indicates a continuous line
      if (!is.na(angle_deg) && (angle_deg <= angle_threshold_low ||
                                angle_deg >= angle_threshold_high)) {
        # Edge is part of a continuous line; do not mark as dangle
        is_continuous = TRUE
        break  # No need to check other edges
      }
    }
    
    # Store min_angle in candidate_edges
    candidate_edges$min_angle_deg[i] = min_angle
    
    # Determine if edge is a dangle
    if (!is_continuous) {
      # Edge is a dangle
      dangle_edge_ids = c(dangle_edge_ids, candidate_edge_id)
      candidate_edges$is_dangle[i] = TRUE
    }
  }
  
  # Merge min_angle_deg and is_dangle back into edges_with_nodes
  edges_with_nodes = edges_with_nodes %>%
    left_join(
      candidate_edges %>% select(edge_id, min_angle_deg, is_dangle),
      by = "edge_id"
    )
  
  # Step 10: Remove Dangle Edges
  cleaned_edges = edges %>%
    filter(!(edge_id %in% dangle_edge_ids))
  
  # Ensure cleaned_edges retains data attributes
  cleaned_edges$edge_id = cleaned_edges$edge_id
  
  # Step 11: Return the Cleaned Network
  cleaned_network = st_transform(cleaned_edges, st_crs(network))
  
  # Ensure cleaned_network is an sf object with data
  cleaned_network = st_sf(cleaned_network)
  
  if (return_debug_data) {
    return(list(
      cleaned_network = cleaned_network,
      edges_with_nodes = edges_with_nodes,
      nodes_df = nodes_df,
      candidate_edges = candidate_edges
    ))
  } else {
    return(cleaned_network)
  }
}

calculate_angle_at_node = function(edge1_id, edge2_id, node_id, edges_with_nodes, nodes_df) {
  # Get edge1
  edge1 = edges_with_nodes %>% filter(edge_id == edge1_id)
  # Get edge2
  edge2 = edges_with_nodes %>% filter(edge_id == edge2_id)
  
  # Coordinates of the node
  node_coords = nodes_df %>% filter(node_id == node_id) %>% select(X, Y) %>% unlist()
  
  # Get the other end of edge1
  if (edge1$from_node_id == node_id) {
    coords1 = c(edge1$x_end, edge1$y_end)
  } else {
    coords1 = c(edge1$x_start, edge1$y_start)
  }
  
  # Get the other end of edge2
  if (edge2$from_node_id == node_id) {
    coords2 = c(edge2$x_end, edge2$y_end)
  } else {
    coords2 = c(edge2$x_start, edge2$y_start)
  }
  
  # Vectors from the node to the other ends
  vector1 = coords1 - node_coords
  vector2 = coords2 - node_coords
  
  # Calculate angle in degrees
  dot_prod = sum(vector1 * vector2)
  mag1 = sqrt(sum(vector1^2))
  mag2 = sqrt(sum(vector2^2))
  cos_theta = dot_prod / (mag1 * mag2)
  # Ensure cos_theta is within [-1, 1] to avoid NaNs
  cos_theta = max(min(cos_theta, 1), -1)
  angle_rad = acos(cos_theta)
  angle_deg = angle_rad * (180 / pi)
  
  # Ensure angle is between 0 and 180 degrees
  if (angle_deg > 180) {
    angle_deg = 360 - angle_deg
  }
  
  return(angle_deg)
}

```
```{r}
library(sf)
library(dplyr)
library(mapview)

# Define the main stem (Edge 1)
line1 = st_linestring(matrix(c(0, -5, 0, 10), ncol = 2, byrow = TRUE))
edge1 = st_sf(edge_id = 1, geometry = st_sfc(line1, crs = 27700))

# Define the left branch (Edge 2)
line2 = st_linestring(matrix(c(0, 10, -5, 10), ncol = 2, byrow = TRUE))
edge2 = st_sf(edge_id = 2, geometry = st_sfc(line2, crs = 27700))

# Define the right branch (Edge 3)
line3 = st_linestring(matrix(c(0, 10, 1, 20), ncol = 2, byrow = TRUE))
edge3 = st_sf(edge_id = 3, geometry = st_sfc(line3, crs = 27700))

# Combine into an sf object
y_network = rbind(edge1, edge2, edge3)
mapview(y_network, color = "black")
```
# latest remove dangles function
```{r}
# Load necessary libraries
library(sf)
library(dplyr)
library(tidyr)
library(igraph)
library(units)
library(lwgeom)
library(mapview)

# Define the remove_dangles function
remove_dangles = function(network, 
                           angle_threshold_low = 30, 
                           angle_threshold_high = 170) {
  
  # Step 0: Assign edge_id if not present
  if(!"edge_id" %in% colnames(network)) {
    network = network %>%
      mutate(edge_id = row_number())
  }
  
  # Step 1: Extract start and end points of each edge
  # Extract start points using st_startpoint
  start_points = st_startpoint(network$geometry) %>%
    st_coordinates() %>%
    as.data.frame() %>%
    rename(X = X, Y = Y) %>%
    mutate(edge_id = network$edge_id,
           type = "start")
  
  # Extract end points using st_endpoint
  end_points = st_endpoint(network$geometry) %>%
    st_coordinates() %>%
    as.data.frame() %>%
    rename(X = X, Y = Y) %>%
    mutate(edge_id = network$edge_id,
           type = "end")
  
  # Combine start and end points
  all_nodes = rbind(start_points, end_points)
  
  # Step 2: Assign unique node IDs based on unique coordinates
  unique_nodes = all_nodes %>%
    distinct(X, Y) %>%
    mutate(node_id = row_number())

  # Step 3: Create an Edge List with From and To Node IDs
  edge_list = all_nodes %>%
    left_join(unique_nodes, by = c("X", "Y")) %>%
    arrange(edge_id, desc(type == "start")) %>%  # Ensure 'start' comes before 'end'
    group_by(edge_id) %>%
    summarize(
      from = first(node_id),  # 'start' node
      to = last(node_id)      # 'end' node
    ) %>%
    ungroup()
  
  # Step 4: Build an igraph Object and Calculate Node Degrees
  g = graph_from_data_frame(edge_list %>% select(from, to), directed = FALSE)
  
  # Compute node degrees
  deg = degree(g)
  
  # Add degrees to unique_nodes
  unique_nodes = unique_nodes %>%
    mutate(degree = deg[node_id])
  
  # Step 5: Identify Candidate Dangles Based on Node Degrees
  dangle_edges = edge_list %>%
    mutate(
      degree_from = unique_nodes$degree[from],
      degree_to = unique_nodes$degree[to],
      is_dangle = (degree_from == 1 | degree_to == 1)
    ) %>%
    filter(is_dangle)
  
  # If no candidate dangles, return the original network
  if(nrow(dangle_edges) == 0) {
    message("No dangles found in the network.")
    return(network)
  }
  
  # Step 6: Define the angle calculation function
  calculate_angle_at_node = function(edge1, edge2, node_id, unique_nodes) {
    # Get coordinates of the common node
    node_coords = c(
      unique_nodes$X[unique_nodes$node_id == node_id],
      unique_nodes$Y[unique_nodes$node_id == node_id]
    )
    
    # Function to get the other end coordinates
    get_other_end = function(edge, node_id, unique_nodes) {
      if (edge$from == node_id) {
        c(
          unique_nodes$X[unique_nodes$node_id == edge$to],
          unique_nodes$Y[unique_nodes$node_id == edge$to]
        )
      } else {
        c(
          unique_nodes$X[unique_nodes$node_id == edge$from],
          unique_nodes$Y[unique_nodes$node_id == edge$from]
        )
      }
    }
    
    # Get the other end coordinates
    coords1 = get_other_end(edge1, node_id, unique_nodes)
    coords2 = get_other_end(edge2, node_id, unique_nodes)
    
    # Create vectors from the common node to the other ends
    vector1 = coords1 - node_coords
    vector2 = coords2 - node_coords
    
    # Calculate dot product and magnitudes
    dot_prod = sum(vector1 * vector2)
    mag1 = sqrt(sum(vector1^2))
    mag2 = sqrt(sum(vector2^2))
    
    # Check for zero-length vectors
    if (mag1 == 0 || mag2 == 0) {
      return(NA)
    }
    
    # Calculate cosine of the angle
    cos_theta = dot_prod / (mag1 * mag2)
    cos_theta = max(min(cos_theta, 1), -1)  # Handle numerical precision
    
    # Calculate angle in degrees
    angle_deg = acos(cos_theta) * (180 / pi)
    
    return(angle_deg)
  }
  
  # Step 7: Identify True Dangles Based on Angle Criteria
  # Initialize a vector to store dangles to remove
  dangles_to_remove = c()
  
  # Iterate over each candidate dangle
  for(i in 1:nrow(dangle_edges)) {
    edge = dangle_edges[i, ]
    
    # Identify the common node (junction node)
    if(edge$degree_from == 1) {
      common_node = edge$to
    } else {
      common_node = edge$from
    }
    
    # Get all other edges connected to the common node
    connected_edges = edge_list %>%
      filter((from == common_node | to == common_node) & edge_id != edge$edge_id)
    
    # Initialize a flag to determine if the edge is a true dangle
    is_true_dangle = TRUE
    
    # Calculate angles between the current edge and all other connected edges
    for(j in 1:nrow(connected_edges)) {
      other_edge = connected_edges[j, ]
      tryCatch({
        angle = calculate_angle_at_node(edge, other_edge, common_node, unique_nodes)
      }, error = function(e) {
        angle = 0
      })

      
      # If angle is NA, skip dangle identification
      if(is.na(angle)) {
        is_true_dangle = FALSE
        break
      }
      
      # Check if the angle is outside the threshold
      if(angle < angle_threshold_low | angle > angle_threshold_high) {
        is_true_dangle = FALSE
        break
      }
    }
    edge$cal_angle = angle
    # If all angles are within the thresholds, mark as dangle
    if(is_true_dangle) {
      dangles_to_remove = c(dangles_to_remove, edge$edge_id)
    }
  }


  # Step 8: Remove Dangles and Return Cleaned Network
  cleaned_network = network %>%
    filter(!edge_id %in% dangles_to_remove)
  
  # Return only the cleaned network
  return(cleaned_network)
}

``` 

```{r}
# Load necessary libraries
library(sf)
library(dplyr)
library(mapview)
library(igraph)

# Define the sample Y-shaped network
# Edge 1: Main Stem
line1 = st_linestring(matrix(c(0, -5, 0, 10), ncol=2, byrow=TRUE))
edge1 = st_sf(edge_id = 1, geometry = st_sfc(line1), crs = 27700)

# Edge 2: Left Branch (Dangle)
line2 = st_linestring(matrix(c(0, 10, -5, 10), ncol=2, byrow=TRUE))
edge2 = st_sf(edge_id = 2, geometry = st_sfc(line2), crs = 27700)

# Edge 3: Right Branch
line3 = st_linestring(matrix(c(0, 10, 1, 20), ncol=2, byrow=TRUE))
edge3 = st_sf(edge_id = 3, geometry = st_sfc(line3), crs = 27700)

# Combine into a single sf object
y_network = rbind(edge1, edge2, edge3)

# Visualize the original network
mapview(y_network, color = "red", layer.name = "Original Network")

# Apply the remove_dangles function
cleaned_network = remove_dangles(network = y_network, 
                                   angle_threshold_low = 30, 
                                   angle_threshold_high = 170)
edin_offroad_net$edge_id = 1:nrow(edin_offroad_net)                                   
cleaned_network_edin = remove_dangles(edin_offroad_net, angle_threshold_low = 30, angle_threshold_high = 170)
mapview(cleaned_network_edin, color = "blue", layer.name = "Cleaned Network")
# View the cleaned network
print("Cleaned Network:")
print(cleaned_network)

# Visualize the cleaned network alongside the original network
mapview(cleaned_network_edin, color = "blue", layer.name = "Cleaned Network") +
  mapview(y_network, color = "red", layer.name = "Original Network", alpha = 0.5)

```

```{r}
# Load necessary libraries
library(sf)
library(dplyr)
library(tidyr)
library(igraph)
library(units)
library(lwgeom)
library(mapview)


# Step 1: Define the network with explicit edge_ids
# Edge 1: Main Stem
line1 = st_linestring(matrix(c(0, -5, 0, 10), ncol=2, byrow=TRUE))
edge1 = st_sf(edge_id = 1, geometry = st_sfc(line1), crs = 27700)

# Edge 2: Left Branch (Dangle)
line2 = st_linestring(matrix(c(0, 10, -5, 10), ncol=2, byrow=TRUE))
edge2 = st_sf(edge_id = 2, geometry = st_sfc(line2), crs = 27700)

# Edge 3: Right Branch
line3 = st_linestring(matrix(c(0, 10, 1, 20), ncol=2, byrow=TRUE))
edge3 = st_sf(edge_id = 3, geometry = st_sfc(line3), crs = 27700)

# Combine into a single sf object
y_network = edin_offroad_net

# Step 2: Extract start and end points of each edge
# Extract start points using st_startpoint
start_points = st_startpoint(y_network$geometry) %>%
  st_coordinates() %>%
  as.data.frame() %>%
  rename(X = X, Y = Y) %>%
  mutate(edge_id = y_network$edge_id,
         type = "start")

# Extract end points using st_endpoint
end_points = st_endpoint(y_network$geometry) %>%
  st_coordinates() %>%
  as.data.frame() %>%
  rename(X = X, Y = Y) %>%
  mutate(edge_id = y_network$edge_id,
         type = "end")

# Combine start and end points
all_nodes = rbind(start_points, end_points)

# Step 3: Assign unique node IDs
# Assign unique node IDs based on unique coordinates
unique_nodes = all_nodes %>%
  distinct(X, Y) %>%
  mutate(node_id = row_number())

# View unique_nodes
print("Unique Nodes:")
print(unique_nodes)

# Step 4: Create an Edge List with From and To Node IDs (Corrected)
edge_list = all_nodes %>%
  left_join(unique_nodes, by = c("X", "Y")) %>%
  arrange(edge_id, desc(type == "start")) %>%  # Ensure 'start' comes before 'end'
  group_by(edge_id) %>%
  summarize(
    from = first(node_id),  # 'start' node
    to = last(node_id)      # 'end' node
  ) %>%
  ungroup()

# View corrected edge_list
print("Edge List:")
print(edge_list)

# Step 5: Build an igraph Object and Calculate Node Degrees
g = graph_from_data_frame(edge_list %>% select(from, to), directed = FALSE)

# Compute node degrees
deg = degree(g)

# Add degrees to unique_nodes
unique_nodes = unique_nodes %>%
  mutate(degree = deg[node_id])

# View nodes with degrees
print("Unique Nodes with Degrees:")
print(unique_nodes)

# Step 6: Identify Candidate Dangles Based on Node Degrees
dangle_edges = edge_list %>%
  mutate(
    degree_from = unique_nodes$degree[from],
    degree_to = unique_nodes$degree[to],
    is_dangle = (degree_from == 1 | degree_to == 1)
  ) %>%
  filter(is_dangle)

# View dangle_edges
print("Candidate Dangle Edges:")
print(dangle_edges)

# Step 7: Calculate Angles to Confirm Dangles
# Corrected function with accurate vector calculations
calculate_angle_at_node = function(edge1, edge2, node_id, unique_nodes) {
  # Get coordinates of the common node
  node_coords = c(
    unique_nodes$X[unique_nodes$node_id == node_id],
    unique_nodes$Y[unique_nodes$node_id == node_id]
  )
  
  # Get the other end coordinates, ensuring we're getting the point away from the junction
  get_other_end = function(edge, node_id, unique_nodes) {
    if (edge$from == node_id) {
      c(
        unique_nodes$X[unique_nodes$node_id == edge$to],
        unique_nodes$Y[unique_nodes$node_id == edge$to]
      )
    } else {
      c(
        unique_nodes$X[unique_nodes$node_id == edge$from],
        unique_nodes$Y[unique_nodes$node_id == edge$from]
      )
    }
  }
  
  # Get vectors from junction to other ends
  coords1 = get_other_end(edge1, node_id, unique_nodes)
  coords2 = get_other_end(edge2, node_id, unique_nodes)
  
  # Create vectors (from junction to endpoints)
  vector1 = coords1 - node_coords
  vector2 = coords2 - node_coords
  
  # Print vectors for verification
  cat("Vector1:", vector1[1], vector1[2], "\n")
  cat("Vector2:", vector2[1], vector2[2], "\n")
  
  # Calculate dot product and magnitudes
  dot_prod = sum(vector1 * vector2)
  mag1 = sqrt(sum(vector1^2))
  mag2 = sqrt(sum(vector2^2))
  
  # Print intermediate calculations
  cat("Dot Product:", dot_prod, "\n")
  cat("Magnitude of Vector1:", mag1, "\n")
  cat("Magnitude of Vector2:", mag2, "\n")
  
  # Check for zero-length vectors
  if (mag1 == 0 || mag2 == 0) {
    cat("Error: Zero-length vector detected.\n")
    return(NA)
  }
  
  # Calculate angle in degrees
  cos_theta = dot_prod / (mag1 * mag2)
  cos_theta = max(min(cos_theta, 1), -1)  # Handle numerical precision
  angle_deg = acos(cos_theta) * (180 / pi)
  
  cat("Calculated Angle:", angle_deg, "degrees\n")
  
  return(angle_deg)
}

# Initialize a vector to store dangles to remove
dangles_to_remove = c()

# Define angle thresholds
angle_threshold_low = 30  # Degrees
angle_threshold_high = 170  # Degrees

# Define the main stem edge_id (e.g., Edge1)
main_stem_edge_id = 1

# Update dangle_edges to exclude the main stem
dangle_edges_filtered = dangle_edges %>%
  filter(edge_id != main_stem_edge_id)

# Iterate over each dangle edge with enhanced debugging
for (i in 1:nrow(dangle_edges_filtered)) {
  edge = dangle_edges_filtered[i, ]
  
  # Identify the node with degree >=3 (common node)
  if (unique_nodes$degree[edge$from] == 1) {
    common_node = edge$to
  } else {
    common_node = edge$from
  }
  
  # Print current edge and common node
  cat("\nProcessing Edge ID:", edge$edge_id, "Common Node ID:", common_node, "\n")
  
  # Get all edges connected to the common node except the current edge
  connected_edges = edge_list %>%
    filter((from == common_node | to == common_node) & edge_id != edge$edge_id)
  
  # Check if there are any connected edges to compare angles
  if (nrow(connected_edges) < 1) {
    cat("No connected edges to compare. Skipping.\n")
    next  # Skip if no other edges are connected
  }
  
  # Calculate angles between the current edge and all other connected edges
  angles = sapply(1:nrow(connected_edges), function(j) {
    other_edge = connected_edges[j, ]
    angle = calculate_angle_at_node(edge, other_edge, common_node, unique_nodes)
    if (!is.na(angle)) {
      cat("Angle with Edge ID:", other_edge$edge_id, "is", angle, "degrees\n")
    } else {
      cat("Angle with Edge ID:", other_edge$edge_id, "could not be calculated.\n")
    }
    return(angle)
  })
  
  # Handle NA angles
  if (any(is.na(angles))) {
    cat("One or more angles are NA. Not marking as dangle.\n")
    next
  }
  
  # Determine if the edge is a dangle based on angles
  # Criteria:
  # - All angles >= angle_threshold_low AND <= angle_threshold_high: dangle
  # - If any angle < angle_threshold_low OR > angle_threshold_high: not a dangle
  if (all(angles >= angle_threshold_low & angles <= angle_threshold_high)) {
    dangles_to_remove = c(dangles_to_remove, edge$edge_id)
    cat("Edge ID:", edge$edge_id, "marked as dangle.\n")
  } else {
    cat("Edge ID:", edge$edge_id, "not marked as dangle due to angle criteria.\n")
  }
}

# View dangles to remove
cat("\nDangles to remove:", dangles_to_remove, "\n")

# Step 8: Remove Dangles and Verify the Cleaned Network
# Remove dangle edges from the network
cleaned_network = y_network %>%
  filter(!edge_id %in% dangles_to_remove)

# View cleaned network
print("Cleaned Network:")
print(cleaned_network)

# Step 9: Visualize the Results
# Visualize the cleaned network and original network
mapview(cleaned_network, color = "blue", layer.name = "Cleaned Network") +
  mapview(y_network, color = "red", layer.name = "Original Network", alpha = 0.5)

```

